### 2024-05-07 article/article.tex
add article structure

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Dummy title} %TODO Please add

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

\author{Paul Hegenberg}{Dummy University Computing Laboratory, [optional: Address], Country \and My second affiliation, Country \and \url{http://www.myhomepage.edu} }{johnqpublic@dummyuni.org}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Joan R. Public\footnote{Optional footnote, e.g. to mark corresponding author}}{Department of Informatics, Dummy College, [optional: Address], Country}{joanrpublic@dummycollege.org}{[orcid]}{[funding]}

\authorrunning{J. Open Access and J.\,R. Public} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

\acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. 
\end{abstract}

\section{Typesetting instructions -- Summary}
\label{sec:typesetting-summary}

LIPIcs is a series of open access high-quality conference proceedings across all fields in informatics established in cooperation with Schloss Dagstuhl. 
In order to do justice to the high scientific quality of the conferences that publish their proceedings in the LIPIcs series, which is ensured by the thorough review process of the respective events, we believe that LIPIcs proceedings must have an attractive and consistent layout matching the standard of the series.
Moreover, the quality of the metadata, the typesetting and the layout must also meet the requirements of other external parties such as indexing service, DOI registry, funding agencies, among others. The guidelines contained in this document serve as the baseline for the authors, editors, and the publisher to create documents that meet as many different requirements as possible.

Please comply with the following instructions when preparing your article for a LIPIcs proceedings volume. 
\paragraph*{Minimum requirements}

\begin{itemize}
\item Use pdflatex and an up-to-date \LaTeX{} system.
\item Use further \LaTeX{} packages and custom made macros carefully and only if required.
\item Use the provided sectioning macros: \verb+\section+, \verb+\subsection+, \verb+\subsubsection+, \linebreak \verb+\paragraph+, \verb+\paragraph*+, and \verb+\subparagraph*+.
\item Provide suitable graphics of at least 300dpi (preferably in PDF format).
\item Use BibTeX and keep the standard style (\verb+plainurl+) for the bibliography.
\item Please try to keep the warnings log as small as possible. Avoid overfull \verb+\hboxes+ and any kind of warnings/errors with the referenced BibTeX entries.
\item Use a spellchecker to correct typos.
\end{itemize}

\paragraph*{Mandatory metadata macros}
Please set the values of the metadata macros carefully since the information parsed from these macros will be passed to publication servers, catalogues and search engines.
Avoid placing macros inside the metadata macros. The following metadata macros/environments are mandatory:
\begin{itemize}
\item \verb+\title+ and, in case of long titles, \verb+\titlerunning+.
\item \verb+\author+, one for each author, even if two or more authors have the same affiliation.
\item \verb+\authorrunning+ and \verb+\Copyright+ (concatenated author names)\\
The \verb+\author+ macros and the \verb+\Copyright+ macro should contain full author names (especially with regard to the first name), while \verb+\authorrunning+ should contain abbreviated first names.
\item \verb+\ccsdesc+ (ACM classification, see \url{https://www.acm.org/publications/class-2012}).
\item \verb+\keywords+ (a comma-separated list of keywords).
\item \verb+\relatedversion+ (if there is a related version, typically the ``full version''); please make sure to provide a persistent URL, e.\,g., at arXiv.
\item \verb+\begin{abstract}...\end{abstract}+ .
\end{itemize}

\paragraph*{Please do not \ldots} %Do not override the \texttt{\seriesstyle}-defaults}
Generally speaking, please do not override the \texttt{lipics-v2021}-style defaults. To be more specific, a short checklist also used by Dagstuhl Publishing during the final typesetting is given below.
In case of \textbf{non-compliance} with these rules Dagstuhl Publishing will remove the corresponding parts of \LaTeX{} code and \textbf{replace it with the \texttt{lipics-v2021} defaults}. In serious cases, we may reject the LaTeX-source and expect the corresponding author to revise the relevant parts.
\begin{itemize}
\item Do not use a different main font. (For example, the \texttt{times} package is forbidden.)
\item Do not alter the spacing of the \texttt{lipics-v2021.cls} style file.
\item Do not use \verb+enumitem+ and \verb+paralist+. (The \texttt{enumerate} package is preloaded, so you can use
 \verb+\begin{enumerate}[(a)]+ or the like.)
\item Do not use ``self-made'' sectioning commands (e.\,g., \verb+\noindent{\bf My+ \verb+Paragraph}+).
\item Do not hide large text blocks using comments or \verb+\iffalse+ $\ldots$ \verb+\fi+ constructions. 
\item Do not use conditional structures to include/exclude content. Instead, please provide only the content that should be published -- in one file -- and nothing else.
\item Do not wrap figures and tables with text. In particular, the package \texttt{wrapfig} is not supported.
\item Do not change the bibliography style. In particular, do not use author-year citations. (The
\texttt{natbib} package is not supported.)
\end{itemize}

\enlargethispage{\baselineskip}

This is only a summary containing the most relevant details. Please read the complete document ``LIPIcs: Instructions for Authors and the \texttt{lipics-v2021} Class'' for all details and don't hesitate to contact Dagstuhl Publishing (\url{mailto:publishing@dagstuhl.de}) in case of questions or comments:
\href{http://drops.dagstuhl.de/styles/lipics-v2021/lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}{\texttt{http://drops.dagstuhl.de/styles/lipics-v2021/\newline lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}}

\section{Lorem ipsum dolor sit amet}

Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}. Praesent convallis orci arcu, eu mollis dolor. Aliquam eleifend suscipit lacinia. Maecenas quam mi, porta ut lacinia sed, convallis ac dui. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse potenti. Donec eget odio et magna ullamcorper vehicula ut vitae libero. Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis. Pellentesque id ante erat. In hac habitasse platea dictumst. Curabitur a scelerisque odio. Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam. Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.

\begin{lemma}[Lorem ipsum]
\label{lemma:lorem}
Vestibulum sodales dolor et dui cursus iaculis. Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper. Proin facilisis gravida rutrum. Etiam sed sollicitudin lorem. Proin pellentesque risus at elit hendrerit pharetra. Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
\end{lemma}

\begin{proof}
Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.

\proofsubparagraph*{Just some paragraph within the proof.}
Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
\begin{claim}
content...
\end{claim}
\begin{claimproof}
content...
    \begin{enumerate}
        \item abc abc abc \claimqedhere{}
    \end{enumerate}
\end{claimproof}

\end{proof}

\begin{corollary}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
\label{lemma:curabitur}
Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum. Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
\end{corollary}

\begin{proposition}\label{prop1}
This is a proposition
\end{proposition}

\autoref{prop1} and \cref{prop1} \ldots

\subsection{Curabitur dictum felis id sapien}

Curabitur dictum \cref{lemma:curabitur} felis id sapien \autoref{lemma:curabitur} mollis ut venenatis tortor feugiat. Curabitur sed velit diam. Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu. Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet. Duis rutrum dapibus ornare. Nulla vehicula vulputate iaculis. Proin a consequat neque. Donec ut rutrum urna. Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum. Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Aenean nec faucibus leo. Cras ut nisl odio, non tincidunt lorem. Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat. Fusce risus libero, convallis at fermentum in, dignissim sed sem. Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}. Donec non suscipit lorem. Nam sit amet enim vitae nisl accumsan pretium.

\begin{lstlisting}[caption={Useless code.},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
for i:=maxint to 0 do 
begin 
    j:=square(root(i));
end;
\end{lstlisting}

\subsection{Proin ac fermentum augue}

Proin ac fermentum augue. Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis. Nulla facilisi. Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac. Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna. Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui. Praesent a pretium quam. Mauris non mauris sit amet eros rutrum aliquam id ut sapien. Nulla aliquet fringilla sagittis. Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor. Nulla cursus aliquet enim. Cras sapien risus, accumsan eu cursus ut, commodo vel velit. Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel. Integer faucibus faucibus felis.

\begin{itemize}
\item Ut vitae diam augue. 
\item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
\item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
\end{itemize}

\begin{remark}
content...
\end{remark}

\section{Pellentesque quis tortor}

Nec urna malesuada sollicitudin. Nulla facilisi. Vivamus aliquam tempus ligula eget ornare. Praesent eget magna ut turpis mattis cursus. Aliquam vel condimentum orci. Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi. Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.

\begin{lemma} [Quisque blandit tempus nunc]
Sed interdum nisl pretium non. Mauris sodales consequat risus vel consectetur. Aliquam erat volutpat. Nunc sed sapien ligula. Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus. Nunc vestibulum nunc ac massa pretium pharetra. Nulla facilisis turpis id augue venenatis blandit. Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
\end{lemma}

Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.

\section{Morbi eros magna}

Morbi eros magna, vestibulum non posuere non, porta eu quam. Maecenas vitae orci risus, eget imperdiet mauris. Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis. Sed condimentum convallis dolor, a dignissim est ultrices eu. Donec consectetur volutpat eros, et ornare dui ultricies id. Vivamus eu augue eget dolor euismod ultrices et sit amet nisi. Vivamus malesuada leo ac leo ullamcorper tempor. Donec justo mi, tempor vitae aliquet non, faucibus eu lacus. Donec dictum gravida neque, non porta turpis imperdiet eget. Curabitur quis euismod ligula.

%%
%% Bibliography
%%

%% Please use bibtex,

\bibliography{lipics-v2021-sample-article}

\appendix

\section{Styles of lists, enumerations, and descriptions}\label{sec:itemStyles}

List of different predefined enumeration styles:

\begin{itemize}
\item \verb|\begin{itemize}...\end{itemize}|
\item \dots
\item \dots
%\item \dots
\end{itemize}

\begin{enumerate}
\item \verb|\begin{enumerate}...\end{enumerate}|
\item \dots
\item \dots
%\item \dots
\end{enumerate}

\begin{alphaenumerate}
\item \verb|\begin{alphaenumerate}...\end{alphaenumerate}|
\item \dots
\item \dots
%\item \dots
\end{alphaenumerate}

\begin{romanenumerate}
\item \verb|\begin{romanenumerate}...\end{romanenumerate}|
\item \dots
\item \dots
%\item \dots
\end{romanenumerate}

\begin{bracketenumerate}
\item \verb|\begin{bracketenumerate}...\end{bracketenumerate}|
\item \dots
\item \dots
%\item \dots
\end{bracketenumerate}

\begin{description}
\item[Description 1] \verb|\begin{description} \item[Description 1]  ...\end{description}|
\item[Description 2] Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\item[Description 3]  \dots
%\item \dots
\end{description}

\cref{testenv-proposition} and \autoref{testenv-proposition} ...

\section{Theorem-like environments}\label{sec:theorem-environments}

List of different predefined enumeration styles:

\begin{theorem}\label{testenv-theorem}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{theorem}

\begin{lemma}\label{testenv-lemma}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{lemma}

\begin{corollary}\label{testenv-corollary}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{corollary}

\begin{proposition}\label{testenv-proposition}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{proposition}

\begin{conjecture}\label{testenv-conjecture}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{conjecture}

\begin{observation}\label{testenv-observation}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{observation}

\begin{exercise}\label{testenv-exercise}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{exercise}

\begin{definition}\label{testenv-definition}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{definition}

\begin{example}\label{testenv-example}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{example}

\begin{note}\label{testenv-note}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{note}

\begin{note*}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{note*}

\begin{remark}\label{testenv-remark}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{remark}

\begin{remark*}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{remark*}

\begin{claim}\label{testenv-claim}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{claim}

\begin{claim*}\label{testenv-claim2}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{claim*}

\begin{proof}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{proof}

\begin{claimproof}
Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
\end{claimproof}

\end{document}

### 2024-06-20 article/article.tex
add evaluation plot and data

@@
 
+\usepackage{pgfplots}
+\pgfplotsset{compat=1.18}
+\usepgfplotslibrary{statistics, units, groupplots}
+
+\pgfplotsset{
+    only if/.style args={entry of #1 is #2}{
+        /pgfplots/boxplot/data filter/.code={
+            \edef\tempa{\thisrow{#1}}
+            \edef\tempb{#2}
+            \ifx\tempa\tempb
+            \else
+                \def\pgfmathresult{}
+            \fi
+        }
+    }
+}
+
+\usepackage{tikz}
+\usetikzlibrary[fadings]
+
+\def\code#1{\texttt{#1}}
+\newcommand\prevsum[0]{\ensuremath{\overset{\leftarrow}{\Sigma}_1}}
+
 %\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
@@
 
-\title{Dummy title} %TODO Please add
+\title{Advanced Data Structures - Succinct Bit Vector}
 
@@
 
-\author{Paul Hegenberg}{Dummy University Computing Laboratory, [optional: Address], Country \and My second affiliation, Country \and \url{http://www.myhomepage.edu} }{johnqpublic@dummyuni.org}{https://orcid.org/0000-0002-1825-0097}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
+\author{Paul Hegenberg}{Karlsruher Institut für Technologie, Germany \and \url{https://blog.hegenberg.dev} }{urlgh@student.kit.edu}{}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
 Please provide at least the name of the affiliation and the country.
@@
 
-\author{Joan R.
-Public\footnote{Optional footnote, e.g.
-to mark corresponding author}}{Department of Informatics, Dummy College, [optional: Address], Country}{joanrpublic@dummycollege.org}{[orcid]}{[funding]}
+\authorrunning{P. Open Access and P. Public} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
 
-\authorrunning{J. Open Access and J.\,R. Public} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'
-
-\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
+\Copyright{Paul Open Access} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/
 
@@
 
-\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords
+\keywords{Bit Vectors} %TODO mandatory; please add comma-separated list of keywords
 
-\category{} %optional, e.g. invited paper
+% \category{} %optional, e.g. invited paper
 
-\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
+% \relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
 %\relatedversiondetails[linktext={opt.
@@
 
-\acknowledgements{I want to thank \dots}%optional
+% \acknowledgements{I want to thank \dots}%optional
 
@@
 \begin{abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Praesent convallis orci arcu, eu mollis dolor.
-Aliquam eleifend suscipit lacinia.
-Maecenas quam mi, porta ut lacinia sed, convallis ac dui.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse potenti.
 \end{abstract}
 
-\section{Typesetting instructions -- Summary}
-\label{sec:typesetting-summary}
+\section{Algorithm}
+In the following section we will present two main data structures which implement the bit vector
+operations
 
-LIPIcs is a series of open access high-quality conference proceedings across all fields in informatics established in cooperation with Schloss Dagstuhl. 
-In order to do justice to the high scientific quality of the conferences that publish their proceedings in the LIPIcs series, which is ensured by the thorough review process of the respective events, we believe that LIPIcs proceedings must have an attractive and consistent layout matching the standard of the series.
-Moreover, the quality of the metadata, the typesetting and the layout must also meet the requirements of other external parties such as indexing service, DOI registry, funding agencies, among others.
-The guidelines contained in this document serve as the baseline for the authors, editors, and the publisher to create documents that meet as many different requirements as possible.
+\subsection{Interleaved}
+The interleaved data structure aims to use the CPU cache as efficiently as possible, where both
+vector information and metadata should fit in the same cache line of 512 bit.
+Our bit vector implementation consists of \code{uint64} blocks, of which we could fit 8 inside one cache line.
+To speed up the computation of rank and select we need metadata, which we store in the first
+64 bits of the line. This data structure is depicted in Figure~\ref{fig:interleaved}. 
+This leaves us with a memory overhead of $12.5\%$. The used metadata $\prevsum$ is an incremental sum of previous bits
+set to one. For an efficient build time of the data structure we make use of the pop count $c(\vec{v})$
+instruction, which counts the number of 1 bits in a subvector $\vec{v}$, in our case an \code{uint64} number.
 
-Please comply with the following instructions when preparing your article for a LIPIcs proceedings volume. 
-\paragraph*{Minimum requirements}
+\begin{figure}[htbp]
+    \centering
+    \tikzfading[name=fade right,
+            left color=transparent!0,
+            right color=transparent!100]
+        \tikzfading[name=fade left,
+            left color=transparent!100,
+            right color=transparent!0]
+    \begin{tikzpicture}
 
-\begin{itemize}
-\item Use pdflatex and an up-to-date \LaTeX{} system.
-\item Use further \LaTeX{} packages and custom made macros carefully and only if required.
-\item Use the provided sectioning macros: \verb+\section+, \verb+\subsection+, \verb+\subsubsection+, \linebreak \verb+\paragraph+, \verb+\paragraph*+, and \verb+\subparagraph*+.
-\item Provide suitable graphics of at least 300dpi (preferably in PDF format).
-\item Use BibTeX and keep the standard style (\verb+plainurl+) for the bibliography.
-\item Please try to keep the warnings log as small as possible. Avoid overfull \verb+\hboxes+ and any kind of warnings/errors with the referenced BibTeX entries.
-\item Use a spellchecker to correct typos.
-\end{itemize}
+        \draw[draw opacity=.4,text opacity=.4,path fading=fade left,densely dotted] (-3,0) rectangle node {$\vec{v}_6$} ++(2,1)
+            ++ (0,-.5) -- ++(1,0);
 
-\paragraph*{Mandatory metadata macros}
-Please set the values of the metadata macros carefully since the information parsed from these macros will be passed to publication servers, catalogues and search engines.
-Avoid placing macros inside the metadata macros. The following metadata macros/environments are mandatory:
-\begin{itemize}
-\item \verb+\title+ and, in case of long titles, \verb+\titlerunning+.
-\item \verb+\author+, one for each author, even if two or more authors have the same affiliation.
-\item \verb+\authorrunning+ and \verb+\Copyright+ (concatenated author names)\\
-The \verb+\author+ macros and the \verb+\Copyright+ macro should contain full author names (especially with regard to the first name), while \verb+\authorrunning+ should contain abbreviated first names.
-\item \verb+\ccsdesc+ (ACM classification, see \url{https://www.acm.org/publications/class-2012}).
-\item \verb+\keywords+ (a comma-separated list of keywords).
-\item \verb+\relatedversion+ (if there is a related version, typically the ``full version''); please make sure to provide a persistent URL, e.\,g., at arXiv.
-\item \verb+\begin{abstract}...\end{abstract}+ .
-\end{itemize}
+        \draw (0,1) rectangle node {$\prevsum$} (2,0) 
+            ++ (0,0) rectangle node {$\vec{v}_0$} ++(2,1)
+            ++ (0,-.5) -- node[label={\dots}] {} ++(2,0)
+            ++ (0,-.5) rectangle node {$\vec{v}_6$} ++(2,1);
 
-\paragraph*{Please do not \ldots} %Do not override the \texttt{\seriesstyle}-defaults}
-Generally speaking, please do not override the \texttt{lipics-v2021}-style defaults.
-To be more specific, a short checklist also used by Dagstuhl Publishing during the final typesetting is given below.
-In case of \textbf{non-compliance} with these rules Dagstuhl Publishing will remove the corresponding parts of \LaTeX{} code and \textbf{replace it with the \texttt{lipics-v2021} defaults}.
-In serious cases, we may reject the LaTeX-source and expect the corresponding author to revise the relevant parts.
-\begin{itemize}
-\item Do not use a different main font. (For example, the \texttt{times} package is forbidden.)
-\item Do not alter the spacing of the \texttt{lipics-v2021.cls} style file.
-\item Do not use \verb+enumitem+ and \verb+paralist+. (The \texttt{enumerate} package is preloaded, so you can use
- \verb+\begin{enumerate}[(a)]+ or the like.)
-\item Do not use ``self-made'' sectioning commands (e.\,g., \verb+\noindent{\bf My+ \verb+Paragraph}+).
-\item Do not hide large text blocks using comments or \verb+\iffalse+ $\ldots$ \verb+\fi+ constructions. 
-\item Do not use conditional structures to include/exclude content. Instead, please provide only the content that should be published -- in one file -- and nothing else.
-\item Do not wrap figures and tables with text. In particular, the package \texttt{wrapfig} is not supported.
-\item Do not change the bibliography style. In particular, do not use author-year citations. (The
-\texttt{natbib} package is not supported.)
-\end{itemize}
+        \draw[|-|] (0,-.3) -- node[label=below:{Vector Line (512 bit)}] {} ++(8,0);
+        \draw[|-|] (0,1.3) -- node[label=above:{64 bit}] {} ++(2,0);
 
-\enlargethispage{\baselineskip}
+        \draw[draw opacity=.4,text opacity=.4,path fading=fade right,densely dotted] (8,.5) -- ++(1,0) 
+            ++(0,-.5) rectangle node {$\prevsum$} ++(2,1);
 
-This is only a summary containing the most relevant details.
-Please read the complete document ``LIPIcs: Instructions for Authors and the \texttt{lipics-v2021} Class'' for all details and don't hesitate to contact Dagstuhl Publishing (\url{mailto:publishing@dagstuhl.de}) in case of questions or comments:
-\href{http://drops.dagstuhl.de/styles/lipics-v2021/lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}{\texttt{http://drops.dagstuhl.de/styles/lipics-v2021/\newline lipics-v2021-authors/lipics-v2021-authors-guidelines.pdf}}
+    \end{tikzpicture}
+    \caption{Interleaved Data Structure}
+    \label{fig:interleaved}    
+\end{figure}
 
-\section{Lorem ipsum dolor sit amet}
+\subsubsection{Rank}
+The \code{rank} operation is defined as $\text{rank}(\alpha, p) \mapsto n$, where $\alpha \in \{0,1\}$
+and $p,n \in \mathbb{N}_0$. Where $n$ is the number of $\alpha$ valued bits set before position $p$.
+To calculate $n$ for a given $\alpha, p$ we first have to calculate the vector line where the Prev. Sum
+is stored. Then we can use bit masks and the pop count operation to count the missing 1 bits.
+$$
+    n' = \prevsum + \sum_{i=0}^{j-1}c(\vec{v}_i) + c(\text{masked}(\vec{v}_j)) 
+$$
+We need to execute 7 pop count operations at most, but since everything fits into one cache line
+we don't need to load additional information from memory. With $\alpha = 1: n = n'$ for $\alpha = 0: n = p - n'$.
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit \cite{DBLP:journals/cacm/Knuth74}.
-Praesent convallis orci arcu, eu mollis dolor.
-Aliquam eleifend suscipit lacinia.
-Maecenas quam mi, porta ut lacinia sed, convallis ac dui.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Suspendisse potenti.
-Donec eget odio et magna ullamcorper vehicula ut vitae libero.
-Maecenas lectus nulla, auctor nec varius ac, ultricies et turpis.
-Pellentesque id ante erat.
-In hac habitasse platea dictumst.
-Curabitur a scelerisque odio.
-Pellentesque elit risus, posuere quis elementum at, pellentesque ut diam.
-Quisque aliquam libero id mi imperdiet quis convallis turpis eleifend.
+\subsubsection{Select}
+The \code{select} operation is defined as $\text{select}(\alpha, n) \mapsto p$, where $p$ is the position
+of the $n$th $\alpha$ valued bit. With the interleaved data structure we use the incremental sum
+to find the vector line with the biggest prev. sum $< n$ using a binary search.
+The following vectors $\vec{v}_i$ will therefore contain the $n$th $\alpha$ bit.
+To find the position we conduct a linear search utilizing pop count, selecting the first
+$\vec{v}_i$ where $c(\vec{v}_i) \geq n$. Next we divide $\vec{v}_i$ into byte sized blocks to apply 
+the same linear search to find the byte containing the $n$th bit.
+To get the position inside the byte we created a lookup table, which is described further in Section~\ref{p:select_lookup}.
 
-\begin{lemma}[Lorem ipsum]
-\label{lemma:lorem}
-Vestibulum sodales dolor et dui cursus iaculis.
-Nullam ullamcorper purus vel turpis lobortis eu tempus lorem semper.
-Proin facilisis gravida rutrum.
-Etiam sed sollicitudin lorem.
-Proin pellentesque risus at elit hendrerit pharetra.
-Integer at turpis varius libero rhoncus fermentum vitae vitae metus.
-\end{lemma}
+\paragraph{Select lookup table}
+\label{p:select_lookup}
+For a fast select lookup we created a lookup table, consisting of 8 columns for each bit,
+as well as 256 rows for each possible byte.
+The $x$-th column in the $y$-th row stores the position of the $x$-th bit in $y$.
+Since most numbers don't contain 8 one bits, we added a validation bit $v$.
+If $v = 1$ the stored position is valid. For each possible byte we need to store 32 bits.
+A schematic view of said lookup table is depicted in Figure~\ref{fig:select_lookup}.
+As this information is constant we included the table as a constant array in the compiled binary.
 
-\begin{proof}
-Cras purus lorem, pulvinar et fermentum sagittis, suscipit quis magna.
+\begin{figure}[htbp]
+    \centering
+    \begin{tikzpicture}
+        \draw[|-|] (0,1.5) -- node[label=above:32 bit] {} ++(9.5,0);
 
-\proofsubparagraph*{Just some paragraph within the proof.}
-Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
-\begin{claim}
-content...
-\end{claim}
-\begin{claimproof}
-content...
-    \begin{enumerate}
-        \item abc abc abc \claimqedhere{}
-    \end{enumerate}
-\end{claimproof}
+        \foreach \x in {1,...,3} \draw (\x*2.25-1.25,1) node {\x};
 
-\end{proof}
+        \draw (7,1) node {$\dots$} ++ (1.5,0) node {8};
 
-\begin{corollary}[Curabitur pulvinar, \cite{DBLP:books/mk/GrayR93}]
-\label{lemma:curabitur}
-Nam liber tempor cum soluta nobis eleifend option congue nihil imperdiet doming id quod mazim placerat facer possim assum.
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna aliquam erat volutpat.
-\end{corollary}
+        \foreach \x in {0,...,1,3}
+        \draw (0,-\x*.75) rectangle node {v} ++(.5,.5) ++(0,-.5) rectangle node {3 bit} ++(1.5,.5)
+            ++(.25,-.5) rectangle node {v} ++(.5,.5) ++(0,-.5) rectangle node {3 bit} ++(1.5,.5)
+            ++(.25,-.5) rectangle node {v} ++(.5,.5) ++(0,-.5) rectangle node {3 bit} ++(1.5,.5)
+            ++(0,-.25) -- node {} ++(1,0)
+            ++(0,-.25) rectangle node {v} ++(.5,.5) ++(0,-.5) rectangle node {3 bit} ++(1.5,.5);
 
-\begin{proposition}\label{prop1}
-This is a proposition
-\end{proposition}
+        % \draw (0,3) rectangle node {v} ++(.5,.5) ++(0,-.5) rectangle node {3 bit} ++(1.5,.5);
 
-\autoref{prop1} and \cref{prop1} \ldots
+        \draw (-.5,.25) node {0} 
+            ++(0,-.75) node {1}
+            ++(0,-.75) node {$\vdots$}
+            ++(0,-.75) node {255};
 
-\subsection{Curabitur dictum felis id sapien}
+    \end{tikzpicture}
+    \caption{Byte select lookup table}
+    \label{fig:select_lookup}
+\end{figure}
 
-Curabitur dictum \cref{lemma:curabitur} felis id sapien \autoref{lemma:curabitur} mollis ut venenatis tortor feugiat.
-Curabitur sed velit diam.
-Integer aliquam, nunc ac egestas lacinia, nibh est vehicula nibh, ac auctor velit tellus non arcu.
-Vestibulum lacinia ipsum vitae nisi ultrices eget gravida turpis laoreet.
-Duis rutrum dapibus ornare.
-Nulla vehicula vulputate iaculis.
-Proin a consequat neque.
-Donec ut rutrum urna.
-Morbi scelerisque turpis sed elit sagittis eu scelerisque quam condimentum.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Aenean nec faucibus leo.
-Cras ut nisl odio, non tincidunt lorem.
-Integer purus ligula, venenatis et convallis lacinia, scelerisque at erat.
-Fusce risus libero, convallis at fermentum in, dignissim sed sem.
-Ut dapibus orci vitae nisl viverra nec adipiscing tortor condimentum \cite{DBLP:journals/cacm/Dijkstra68a}.
-Donec non suscipit lorem.
-Nam sit amet enim vitae nisl accumsan pretium.
+\section{Evolution}
 
-\begin{lstlisting}[caption={Useless code.},label=list:8-6,captionpos=t,float,abovecaptionskip=-\medskipamount]
-for i:=maxint to 0 do 
-begin 
-    j:=square(root(i));
-end;
-\end{lstlisting}
+\section{Evaluation}
+In this chapter we will evaluate our predicted runtime against measurements.
+For the pre-computation we calculate an incremental sum $\prevsum$ every 7 blocks of 64 bits.
+This should scale linearly with the vector size $|\vec{V}|=n$.
+As for \code{rank} we can calculate the position of the vector line in constant time.
+For \code{select} we need to first run a binary search in $\mathcal{O}(\log{n})$ to find the vector line.
+After that we compute a linear search in $\mathcal{O}(7)$ as there is a maximum of 7 blocks we need to search.
+The runtime will be dominated by the binary search after a certain time.
+In Figure~\ref{fig:performance_eval} we can validate these predictions.
 
-\subsection{Proin ac fermentum augue}
+\begin{figure}[htbp]
+    \centering
+    \begin{tikzpicture}    
+        \begin{axis}[
+            legend columns=3,
+            width=\textwidth*0.9,
+            height=7cm,
+            ymode=log,
+            log basis y=10,
+            ylabel={runtime},
+            xlabel={vector size},
+            y SI prefix=milli,
+            y unit=s,
+            every axis legend/.append style={
+                at={(0.5,1.03)},
+                anchor=south
+            }, 
+            xmin=28,
+            xticklabel={$2^{\pgfmathparse{\tick}\pgfmathprintnumber{\pgfmathresult}}$},
+            enlarge x limits=auto,
+            grid=both,
+            tick align=outside,
+            tickpos=left,
+        ]
+            \addplot [
+                mark=+,
+                error bars/.cd,
+                    y dir=both,
+                    y explicit,
+            ] table [x=bits,y=mean,y error=error] {errorplot.dat};
+            \addlegendentry{Runtime}
 
-Proin ac fermentum augue.
-Nullam bibendum enim sollicitudin tellus egestas lacinia euismod orci mollis.
-Nulla facilisi.
-Vivamus volutpat venenatis sapien, vitae feugiat arcu fringilla ac.
-Mauris sapien tortor, sagittis eget auctor at, vulputate pharetra magna.
-Sed congue, dui nec vulputate convallis, sem nunc adipiscing dui, vel venenatis mauris sem in dui.
-Praesent a pretium quam.
-Mauris non mauris sit amet eros rutrum aliquam id ut sapien.
-Nulla aliquet fringilla sagittis.
-Pellentesque eu metus posuere nunc tincidunt dignissim in tempor dolor.
-Nulla cursus aliquet enim.
-Cras sapien risus, accumsan eu cursus ut, commodo vel velit.
-Praesent aliquet consectetur ligula, vitae iaculis ligula interdum vel.
-Integer faucibus faucibus felis.
+            \addplot [
+                mark=+,
+                color=red,
+                error bars/.cd,
+                    y dir=both,
+                    y explicit,
+            ] table [x=bits,y=commandMean,y error=commandError] {errorplot.dat};
+            \addlegendentry{Command Execution}
 
-\begin{itemize}
-\item Ut vitae diam augue. 
-\item Integer lacus ante, pellentesque sed sollicitudin et, pulvinar adipiscing sem. 
-\item Maecenas facilisis, leo quis tincidunt egestas, magna ipsum condimentum orci, vitae facilisis nibh turpis et elit. 
-\end{itemize}
+            \addplot [
+                mark=+,
+                color=blue,
+                error bars/.cd,
+                    y dir=both,
+                    y explicit,
+            ] table [x=bits,y=precomMean,y error=precomError] {errorplot.dat};
+            \addlegendentry{Precomputation}
+        \end{axis} 
+    \end{tikzpicture}
+    \caption{Performance Evaluation of an increasing vector size and a constant 100 000 commands.}
+    \label{fig:performance_eval}
+\end{figure}
 
-\begin{remark}
-content...
-\end{remark}
+% \begin{tikzpicture}
+% 	\begin{axis}[
+%         boxplot,
+% 		boxplot/draw direction=y,
+% 		ylabel={runtime},
+%         y SI prefix=milli,y unit=s,
+%         xmin=0,
+%         xticklabel={$2^{\pgfmathparse{8+\tick}\pgfmathprintnumber{\pgfmathresult}}$},
+%         % x tick label as interval,
+%         xlabel={vector length},
+% 	]
 
-\section{Pellentesque quis tortor}
+%     % \addplot table[y=time, only if={entry of bits is 31}]{benchmark.dat}; 
+%     % \addplot table[y=time, only if={entry of bits is 32}]{benchmark.dat}; 
+%     \pgfplotsinvokeforeach{8,...,32}{ 
+%         \addplot table[y=time,x=bits,only if={entry of bits is #1}]{benchmark.dat}; 
+%         % \addlegendentry{Cond. $2^{#1}$};
+%     }
 
-Nec urna malesuada sollicitudin.
-Nulla facilisi.
-Vivamus aliquam tempus ligula eget ornare.
-Praesent eget magna ut turpis mattis cursus.
-Aliquam vel condimentum orci.
-Nunc congue, libero in gravida convallis \cite{DBLP:conf/focs/HopcroftPV75}, orci nibh sodales quam, id egestas felis mi nec nisi.
-Suspendisse tincidunt, est ac vestibulum posuere, justo odio bibendum urna, rutrum bibendum dolor sem nec tellus.
+% 	\end{axis}
+%     \begin{axis}[
+%         boxplot,
+% 		boxplot/draw direction=y,
+% 		ylabel={runtime},
+%         y SI prefix=milli,y unit=s,
+%         xmin=0,
+%         xticklabel={$2^{\pgfmathparse{8+\tick}\pgfmathprintnumber{\pgfmathresult}}$},
+%         % x tick label as interval,
+%         xlabel={vector length},
+% 	]
 
-\begin{lemma} [Quisque blandit tempus nunc]
-Sed interdum nisl pretium non.
-Mauris sodales consequat risus vel consectetur.
-Aliquam erat volutpat.
-Nunc sed sapien ligula.
-Proin faucibus sapien luctus nisl feugiat convallis faucibus elit cursus.
-Nunc vestibulum nunc ac massa pretium pharetra.
-Nulla facilisis turpis id augue venenatis blandit.
-Cum sociis natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus.
-\end{lemma}
+%     % \addplot table[y=time, only if={entry of bits is 31}]{benchmark.dat}; 
+%     % \addplot table[y=time, only if={entry of bits is 32}]{benchmark.dat}; 
+%     \pgfplotsinvokeforeach{8,...,32}{ 
+%         \addplot table[y=time,x=bits,only if={entry of bits is #1}]{benchmark.dat}; 
+%         % \addlegendentry{Cond. $2^{#1}$};
+%      }
 
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-
-\section{Morbi eros magna}
-
-Morbi eros magna, vestibulum non posuere non, porta eu quam.
-Maecenas vitae orci risus, eget imperdiet mauris.
-Donec massa mauris, pellentesque vel lobortis eu, molestie ac turpis.
-Sed condimentum convallis dolor, a dignissim est ultrices eu.
-Donec consectetur volutpat eros, et ornare dui ultricies id.
-Vivamus eu augue eget dolor euismod ultrices et sit amet nisi.
-Vivamus malesuada leo ac leo ullamcorper tempor.
-Donec justo mi, tempor vitae aliquet non, faucibus eu lacus.
-Donec dictum gravida neque, non porta turpis imperdiet eget.
-Curabitur quis euismod ligula.
+%     \end{axis}
+% \end{tikzpicture}
 
@@
 
-\section{Styles of lists, enumerations, and descriptions}\label{sec:itemStyles}
-
-List of different predefined enumeration styles:
-
-\begin{itemize}
-\item \verb|\begin{itemize}...\end{itemize}|
-\item \dots
-\item \dots
-%\item \dots
-\end{itemize}
-
-\begin{enumerate}
-\item \verb|\begin{enumerate}...\end{enumerate}|
-\item \dots
-\item \dots
-%\item \dots
-\end{enumerate}
-
-\begin{alphaenumerate}
-\item \verb|\begin{alphaenumerate}...\end{alphaenumerate}|
-\item \dots
-\item \dots
-%\item \dots
-\end{alphaenumerate}
-
-\begin{romanenumerate}
-\item \verb|\begin{romanenumerate}...\end{romanenumerate}|
-\item \dots
-\item \dots
-%\item \dots
-\end{romanenumerate}
-
-\begin{bracketenumerate}
-\item \verb|\begin{bracketenumerate}...\end{bracketenumerate}|
-\item \dots
-\item \dots
-%\item \dots
-\end{bracketenumerate}
-
-\begin{description}
-\item[Description 1] \verb|\begin{description} \item[Description 1]  ...\end{description}|
-\item[Description 2] Fusce eu leo nisi.
-Cras eget orci neque, eleifend dapibus felis.
-Duis et leo dui.
-Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\item[Description 3]  \dots
-%\item \dots
-\end{description}
-
-\cref{testenv-proposition} and \autoref{testenv-proposition} ...
-
-\section{Theorem-like environments}\label{sec:theorem-environments}
-
-List of different predefined enumeration styles:
-
-\begin{theorem}\label{testenv-theorem}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{theorem}
-
-\begin{lemma}\label{testenv-lemma}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{lemma}
-
-\begin{corollary}\label{testenv-corollary}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{corollary}
-
-\begin{proposition}\label{testenv-proposition}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{proposition}
-
-\begin{conjecture}\label{testenv-conjecture}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{conjecture}
-
-\begin{observation}\label{testenv-observation}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{observation}
-
-\begin{exercise}\label{testenv-exercise}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{exercise}
-
-\begin{definition}\label{testenv-definition}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{definition}
-
-\begin{example}\label{testenv-example}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{example}
-
-\begin{note}\label{testenv-note}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{note}
-
-\begin{note*}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{note*}
-
-\begin{remark}\label{testenv-remark}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{remark}
-
-\begin{remark*}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{remark*}
-
-\begin{claim}\label{testenv-claim}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{claim}
-
-\begin{claim*}\label{testenv-claim2}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{claim*}
-
-\begin{proof}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{proof}
-
-\begin{claimproof}
-Fusce eu leo nisi. Cras eget orci neque, eleifend dapibus felis. Duis et leo dui. Nam vulputate, velit et laoreet porttitor, quam arcu facilisis dui, sed malesuada risus massa sit amet neque.
-\end{claimproof}
-
 \end{document}

### 2024-07-01 article/article.tex
add second benchmark

@@
 
-%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
+% \pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
+\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
+\nolinenumbers
 
@@
 \usepgfplotslibrary{statistics, units, groupplots}
+
+\usepackage{subcaption}
 
@@
 
-\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm
+\ccsdesc[100]{Theory of computation~Data structures design and analysis} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm
 
@@
 
-%TODO mandatory: add short abstract of the document
 \begin{abstract}
+We introduce an interleaved data structure optimized for CPU cache efficiency to implement \code{rank},
+\code{select}, and \code{access} operations. This structure stores vector data and metadata within a
+512-bit cache line, using 64-bit blocks. Metadata, as an incremental sum of set bits, supports fast 
+rank and select operations with a 12.5\% memory overhead.
+The \code{rank} operation achieves constant execution time, by using incremental sums and pop counts.
+The \code{select} operation locates the nth bit occurrence via binary search on metadata and linear search,
+aided by a byte-level lookup table.
+Performance evaluation confirms that pre-computation scales linearly with vector size and \code{select}
+execution scales logarithmic, with variations in speed depending on CPU architecture.
 \end{abstract}
@@
 \section{Algorithm}
-In the following section we will present two main data structures which implement the bit vector
-operations
-
-\subsection{Interleaved}
+In the following section we will present an interleaved datastructure
+that implements \code{rank}, \code{select} and \code{access} operations.
 The interleaved data structure aims to use the CPU cache as efficiently as possible, where both
 vector information and metadata should fit in the same cache line of 512 bit.
-Our bit vector implementation consists of \code{uint64} blocks, of which we could fit 8 inside one cache line.
+Our bit vector implementation consists of \code{uint64} blocks, of which we can fit 8 inside one cache line
+$L_i = (\prevsum, \vec{v}_0, \dots, \vec{v}_6)$.
 To speed up the computation of rank and select we need metadata, which we store in the first
-64 bits of the line. This data structure is depicted in Figure~\ref{fig:interleaved}. 
+64 bits of the line.
+This data structure is depicted in Figure~\ref{fig:interleaved}. 
 This leaves us with a memory overhead of $12.5\%$. The used metadata $\prevsum$ is an incremental sum of previous bits
@@
 
-\subsubsection{Rank}
+\subsection{Rank}
 The \code{rank} operation is defined as $\text{rank}(\alpha, p) \mapsto n$, where $\alpha \in \{0,1\}$
 and $p,n \in \mathbb{N}_0$. Where $n$ is the number of $\alpha$ valued bits set before position $p$.
-To calculate $n$ for a given $\alpha, p$ we first have to calculate the vector line where the Prev. Sum
-is stored. Then we can use bit masks and the pop count operation to count the missing 1 bits.
+To calculate $n$ for a given $\alpha, p$ we first have to find the vector line $L_i$ where the
+previous sum $\prevsum$ is stored, by calculating $i=\frac{p}{64 \cdot 7}$.
+Using $j = \frac{p}{64} \mod 7$ we then know how many subvectors $\vec{v}$ we need to add to $\prevsum$.
+Then we can use bit masks and the pop count operation to count the missing 1 bits.
 $$
-    n' = \prevsum + \sum_{i=0}^{j-1}c(\vec{v}_i) + c(\text{masked}(\vec{v}_j)) 
+    n' = \prevsum + \sum_{s=0}^{j-1}c(\vec{v}_s) + c(\text{masked}_{m}(\vec{v}_j))\\
 $$
-We need to execute 7 pop count operations at most, but since everything fits into one cache line
-we don't need to load additional information from memory. With $\alpha = 1: n = n'$ for $\alpha = 0: n = p - n'$.
+Where $m = p \mod 64$, With $\alpha = 1: n = n'$ for $\alpha = 0: n = p - n'$.
+Considering the runtime, we need to execute 7 pop count operations at most, since $0 \leq j \leq 6$.
+Resulting in a constant execution time.
+Since everything fits into one cache line we don't need to load additional information from memory,
+which should improve performance.
 
-\subsubsection{Select}
-The \code{select} operation is defined as $\text{select}(\alpha, n) \mapsto p$, where $p$ is the position
+\subsection{Select}
+The \code{select} operation is defined as $\text{select}(\alpha, n) \mapsto p$, where $n \in \mathbb{N}$ and $p$ is the position
 of the $n$th $\alpha$ valued bit. With the interleaved data structure we use the incremental sum
-to find the vector line with the biggest prev. sum $< n$ using a binary search.
-The following vectors $\vec{v}_i$ will therefore contain the $n$th $\alpha$ bit.
+to find the vector line with the biggest prev. sum $\prevsum < n$ using a binary search.
+The following subvectors $\vec{v}_i$ will therefore contain the $n$th $\alpha$ bit.
 To find the position we conduct a linear search utilizing pop count, selecting the first
@@
 
-\paragraph{Select lookup table}
+\subsubsection{Select lookup table}
 \label{p:select_lookup}
@@
 
-\section{Evolution}
-
 \section{Evaluation}
@@
 After that we compute a linear search in $\mathcal{O}(7)$ as there is a maximum of 7 blocks we need to search.
-The runtime will be dominated by the binary search after a certain time.
-In Figure~\ref{fig:performance_eval} we can validate these predictions.
-
+The runtime should be dominated by the binary search.
+To validate these predictions we conducted an experiment were a bitvector $\vec{V}$ with increasing
+size was randomly generated as well as a fixed amount of $100 000$ commands.
+We then measured the runtime for the pre-computation as well as the command execution.
+In Figure~\ref{fig:performance_eval} the recorded data is visualized.
 \begin{figure}[htbp]
     \centering
-    \begin{tikzpicture}    
-        \begin{axis}[
-            legend columns=3,
-            width=\textwidth*0.9,
-            height=7cm,
-            ymode=log,
-            log basis y=10,
-            ylabel={runtime},
-            xlabel={vector size},
-            y SI prefix=milli,
-            y unit=s,
-            every axis legend/.append style={
-                at={(0.5,1.03)},
-                anchor=south
-            }, 
-            xmin=28,
-            xticklabel={$2^{\pgfmathparse{\tick}\pgfmathprintnumber{\pgfmathresult}}$},
-            enlarge x limits=auto,
-            grid=both,
-            tick align=outside,
-            tickpos=left,
-        ]
-            \addplot [
-                mark=+,
-                error bars/.cd,
-                    y dir=both,
-                    y explicit,
-            ] table [x=bits,y=mean,y error=error] {errorplot.dat};
-            \addlegendentry{Runtime}
-
-            \addplot [
-                mark=+,
-                color=red,
-                error bars/.cd,
-                    y dir=both,
-                    y explicit,
-            ] table [x=bits,y=commandMean,y error=commandError] {errorplot.dat};
-            \addlegendentry{Command Execution}
-
-            \addplot [
-                mark=+,
-                color=blue,
-                error bars/.cd,
-                    y dir=both,
-                    y explicit,
-            ] table [x=bits,y=precomMean,y error=precomError] {errorplot.dat};
-            \addlegendentry{Precomputation}
-        \end{axis} 
-    \end{tikzpicture}
-    \caption{Performance Evaluation of an increasing vector size and a constant 100 000 commands.}
+    \input{images/benchmark.tex}
+    \caption{%
+        Performance Evaluation of an increasing vector size and a constant 100 000 commands.
+        The benchmark was run on both:
+        a 12th Gen Intel® Core™ i5-1240P × 16 with 32 GiB RAM, 1 TB SSD running Fedora Linux 40
+        and on a dedicated vCPU Server (Hetzner CCX33) with 
+        8 Cores, 32 GiB RAM and 240 GiB SSD running on AMD Milan EPYC™ 7003 or AMD Genoa EPYC™ 9654
+        running Ubuntu 24.04.
+    }
     \label{fig:performance_eval}
 \end{figure}
-
-% \begin{tikzpicture}
-% 	\begin{axis}[
-%         boxplot,
-% 		boxplot/draw direction=y,
-% 		ylabel={runtime},
-%         y SI prefix=milli,y unit=s,
-%         xmin=0,
-%         xticklabel={$2^{\pgfmathparse{8+\tick}\pgfmathprintnumber{\pgfmathresult}}$},
-%         % x tick label as interval,
-%         xlabel={vector length},
-% 	]
-
-%     % \addplot table[y=time, only if={entry of bits is 31}]{benchmark.dat}; 
-%     % \addplot table[y=time, only if={entry of bits is 32}]{benchmark.dat}; 
-%     \pgfplotsinvokeforeach{8,...,32}{ 
-%         \addplot table[y=time,x=bits,only if={entry of bits is #1}]{benchmark.dat}; 
-%         % \addlegendentry{Cond. $2^{#1}$};
-%     }
-
-% 	\end{axis}
-%     \begin{axis}[
-%         boxplot,
-% 		boxplot/draw direction=y,
-% 		ylabel={runtime},
-%         y SI prefix=milli,y unit=s,
-%         xmin=0,
-%         xticklabel={$2^{\pgfmathparse{8+\tick}\pgfmathprintnumber{\pgfmathresult}}$},
-%         % x tick label as interval,
-%         xlabel={vector length},
-% 	]
-
-%     % \addplot table[y=time, only if={entry of bits is 31}]{benchmark.dat}; 
-%     % \addplot table[y=time, only if={entry of bits is 32}]{benchmark.dat}; 
-%     \pgfplotsinvokeforeach{8,...,32}{ 
-%         \addplot table[y=time,x=bits,only if={entry of bits is #1}]{benchmark.dat}; 
-%         % \addlegendentry{Cond. $2^{#1}$};
-%      }
-
-%     \end{axis}
-% \end{tikzpicture}
+Analyzing our data we can confirm the predicted runtime, where the pre-computation scales linearly
+with the vector length, and the command execution time scaling logarithmic.
+We also notice that the pre-computation on the AMD setup being faster while the command execution being
+slower in comparison to the Intel setup.
 
@@
 
-\bibliography{lipics-v2021-sample-article}
+% \bibliography{lipics-v2021-sample-article}
 

### 2024-07-01 article/images/benchmark.tex
add second benchmark

\begin{tikzpicture}
    \begin{groupplot}[group style={group size= 2 by 1},height=6cm,width=.45\textwidth]
        \nextgroupplot[
            legend columns=-1,
            legend to name=usageLegend,
            ymin=1,
            ymax=1000,
            ymode=log,
            log basis y=10,
            ylabel={runtime},
            xlabel={vector size},
            y SI prefix=milli,
            y unit=s,
            xmin=27,
            xticklabel={$2^{\pgfmathparse{\tick}\pgfmathprintnumber{\pgfmathresult}}$},
            enlarge x limits=auto,
            grid=both,
            tick align=outside,
            tickpos=left,
            title=Intel,
        ]
        \addplot [
            mark=+,
            error bars/.cd,
                y dir=both,
                y explicit,
        ] table [x=bits,y=mean,y error=error] {benchmarks/framework.dat};
        \addlegendentry{Runtime}

        \addplot [
            mark=+,
            color=red,
            error bars/.cd,
                y dir=both,
                y explicit,
        ] table [x=bits,y=commandMean,y error=commandError] {benchmarks/framework.dat};
        \addlegendentry{Command Execution}

        \addplot [
            mark=+,
            color=blue,
            error bars/.cd,
                y dir=both,
                y explicit,
        ] table [x=bits,y=precomMean,y error=precomError] {benchmarks/framework.dat};
        \addlegendentry{Precomputation}

        \coordinate (topleft) at (rel axis cs:0,1);% coordinate at top of the first plot

        \nextgroupplot[
            ymode=log,
            ymin=1,
            ymax=1000,
            log basis y=10,
            xlabel={vector size},
            xmin=27,
            yticklabel=\empty,
            xticklabel={$2^{\pgfmathparse{\tick}\pgfmathprintnumber{\pgfmathresult}}$},
            enlarge x limits=auto,
            grid=both,
            tick align=outside,
            tickpos=left,
            title=AMD,
        ]

        \addplot [
                mark=+,
                color=black,
                error bars/.cd,
                    y dir=both,
                    y explicit,
            ] table [x=bits,y=mean,y error=error] {benchmarks/hetzner.dat};

            \addplot [
                mark=+,
                color=red,
                error bars/.cd,
                    y dir=both,
                    y explicit,
            ] table [x=bits,y=commandMean,y error=commandError] {benchmarks/hetzner.dat};

            \addplot [
                mark=+,
                color=blue,
                error bars/.cd,
                    y dir=both,
                    y explicit,
            ] table [x=bits,y=precomMean,y error=precomError] {benchmarks/hetzner.dat};

        \coordinate (topright) at (rel axis cs:1,1);% coordinate at top of the second plot

    \end{groupplot}
    \path (topleft)--(topright) coordinate[midway] (group center);
    \node[align=center,above,yshift=.8cm] at(group center) {\pgfplotslegendfromname{usageLegend}};
\end{tikzpicture}

### 2024-07-03 article/article.tex
add algorithm implementation

@@
 \hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-\nolinenumbers
+% \nolinenumbers
 
@@
 \section{Algorithm}
-In the following section we will present an interleaved datastructure
+In the following section we will present a datastructure
 that implements \code{rank}, \code{select} and \code{access} operations.
-The interleaved data structure aims to use the CPU cache as efficiently as possible, where both
-vector information and metadata should fit in the same cache line of 512 bit.
+The data structure aims to use the CPU cache as efficiently as possible, by interleaving both
+vector information and metadata to fit in the same cache line of 512 bit.
 Our bit vector implementation consists of \code{uint64} blocks, of which we can fit 8 inside one cache line
@@
 In Figure~\ref{fig:performance_eval} the recorded data is visualized.
-\begin{figure}[htbp]
+\begin{figure}[t]
     \centering
@@
         and on a dedicated vCPU Server (Hetzner CCX33) with 
-        8 Cores, 32 GiB RAM and 240 GiB SSD running on AMD Milan EPYC™ 7003 or AMD Genoa EPYC™ 9654
-        running Ubuntu 24.04.
+        8 Cores, 32 GiB RAM and 240 GiB SSD running on AMD Milan EPYC™ 7003 running Ubuntu 24.04.
     }

### 2024-07-03 benchmark/generate.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+
+for i in {8..34}
+do
+    vecBlocks64=$((2**(i-6)))
+    mkdir ./"${i}"
+   ../generator -l ${vecBlocks64} -c 1000000 -o ./"${i}"
+done
+

### 2024-07-03 benchmark/run.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+echo -e "New run" >> results.txt
+for i in {8..34}
+do
+    for j in {0..10}
+    do
+        ../bitvector -verbose ./"${i}"/commands.txt ./"${i}"/out.txt >> results.txt
+        echo -e " bits=${i}" >> results.txt
+    done
+done

### 2024-07-03 build.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+go mod download
+
+go build -o bitvector ./cmd/bitvector/main.go

### 2024-07-03 build_generator.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+go mod download
+
+go build -o generator ./cmd/bitvector_generator/main.go

### 2024-07-03 cmd/bitvector/main.go

@@ -1,0 +1,54 @@
+package main
+
+import (
+	"flag"
+	"log"
+	"os"
+	"runtime/pprof"
+
+	"github.com/paulheg/kit_advanced_data_structures/internal/bitvector"
+)
+
+var cpuprofile = flag.String("cpuprofile", "", "write cpu profile to file")
+var verbose = flag.Bool("verbose", false, "add more parameters to the output")
+
+func main() {
+
+	flag.Parse()
+
+	// if we want to record a CPU profile an output filepath will be set
+	if *cpuprofile != "" {
+		f, err := os.Create(*cpuprofile)
+		if err != nil {
+			log.Fatal(err)
+		}
+		pprof.StartCPUProfile(f)
+		defer pprof.StopCPUProfile()
+	}
+
+	// parse the commandline args these are the input and output paths
+	files := flag.Args()
+	if len(files) != 2 {
+		log.Fatal("wrong number of arguments, need [input] [output]")
+	}
+
+	inputPath := files[0]
+	inputFile, err := os.Open(inputPath)
+	if err != nil {
+		log.Fatal("could not open input file:", err)
+	}
+	defer inputFile.Close()
+
+	outputPath := files[1]
+	outputFile, err := os.Create(outputPath)
+	if err != nil {
+		log.Fatal("could not create output file:", err)
+	}
+	defer outputFile.Close()
+
+	// here the actual processing begins
+	err = bitvector.ProcessFile(inputFile, outputFile, os.Stdout, *verbose)
+	if err != nil {
+		log.Fatal("error processing file:", err)
+	}
+}

### 2024-07-03 cmd/bitvector_generator/main.go

@@ -1,0 +1,68 @@
+package main
+
+import (
+	"log"
+	"os"
+	"path"
+
+	"github.com/paulheg/kit_advanced_data_structures/internal/bitvector"
+	"github.com/urfave/cli/v2"
+)
+
+func main() {
+
+	app := cli.App{
+		Name: "generator",
+		Flags: []cli.Flag{
+			&cli.PathFlag{
+				Name: "output-dir",
+				Aliases: []string{
+					"o", "output",
+				},
+				Value: ".",
+			},
+			&cli.Uint64Flag{
+				Name: "vector-length",
+				Aliases: []string{
+					"l", "length",
+				},
+				Value: 10,
+			},
+			&cli.Uint64Flag{
+				Name: "commands",
+				Aliases: []string{
+					"n", "c",
+				},
+				Value: 10,
+			},
+		},
+		Action: func(ctx *cli.Context) error {
+
+			outputPath := ctx.Path("output-dir")
+
+			fExpected, err := os.Create(path.Join(outputPath, "expected.txt"))
+			if err != nil {
+				return err
+			}
+			defer fExpected.Close()
+
+			fCommands, err := os.Create(path.Join(outputPath, "commands.txt"))
+			if err != nil {
+				return err
+			}
+			defer fCommands.Close()
+
+			err = bitvector.GenerateRandomTestCase(ctx.Uint64("vector-length"), ctx.Uint64("commands"), fCommands, fExpected)
+			if err != nil {
+				return err
+			}
+
+			log.Println("finished...")
+			return nil
+		},
+	}
+
+	if err := app.Run(os.Args); err != nil {
+		log.Fatal(err)
+	}
+}

### 2024-07-03 internal/bitvector/command.go

@@ -1,0 +1,85 @@
+package bitvector
+
+import (
+	"errors"
+	"fmt"
+	"strconv"
+
+	"github.com/paulheg/kit_advanced_data_structures/pkg/bit"
+)
+
+type Command string
+
+const (
+	Access Command = "access"
+	Rank   Command = "rank"
+	Select Command = "select"
+)
+
+var Commands []Command = []Command{
+	Access, Rank, Select,
+}
+
+type CommandFuncGenerator func(args []string) (CommandFunc, error)
+type CommandFunc func(vec bit.RankSelectVector) (uint64, error)
+
+var CommandExecutors = map[Command]CommandFuncGenerator{
+	Access: func(args []string) (CommandFunc, error) {
+		if len(args) != 1 {
+			return nil, errors.New("access only accepts one argument")
+		}
+
+		position, err := strconv.ParseUint(args[0], 10, 64)
+		if err != nil {
+			return nil, fmt.Errorf("position argument not a valid number: %w", err)
+		}
+
+		return func(vec bit.RankSelectVector) (uint64, error) {
+			b := vec.Access(position)
+			v := uint64(0)
+			if b {
+				v = 1
+			}
+			return v, nil
+		}, nil
+	},
+	Rank: func(args []string) (CommandFunc, error) {
+		if len(args) != 2 {
+			return nil, errors.New("rank only accepts two arguments")
+		}
+
+		alpha, err := strconv.ParseBool(args[0])
+		if err != nil {
+			return nil, fmt.Errorf("alpha argument not valid: %w", err)
+		}
+
+		position, err := strconv.ParseUint(args[1], 10, 64)
+		if err != nil {
+			return nil, fmt.Errorf("position argument not valid: %w", err)
+		}
+
+		return func(vec bit.RankSelectVector) (uint64, error) {
+			return vec.Rank(alpha, position), nil
+		}, nil
+
+	},
+	Select: func(args []string) (CommandFunc, error) {
+		if len(args) != 2 {
+			return nil, errors.New("select only accepts two arguments")
+		}
+
+		alpha, err := strconv.ParseBool(args[0])
+		if err != nil {
+			return nil, fmt.Errorf("alpha argument not valid: %w", err)
+		}
+
+		position, err := strconv.ParseUint(args[1], 10, 64)
+		if err != nil {
+			return nil, fmt.Errorf("position argument not valid: %w", err)
+		}
+
[…]

### 2024-07-03 internal/bitvector/file_processor.go

@@ -1,0 +1,129 @@
+package bitvector
+
+import (
+	"bufio"
+	"errors"
+	"fmt"
+	"io"
+	"strconv"
+	"strings"
+	"time"
+
+	"github.com/paulheg/kit_advanced_data_structures/pkg/bit"
+)
+
+func ProcessFile(input io.Reader, output io.Writer, statOut io.Writer, verbose bool) error {
+
+	var err error
+
+	reader := bufio.NewReader(input)
+
+	noOfCommands := 0
+	line, err := reader.ReadString('\n')
+	if err != nil {
+		return fmt.Errorf("number of commands could not be read: %w", err)
+	} else {
+		noOfCommands, err = strconv.Atoi(line[:len(line)-1])
+		if err != nil {
+			return errors.New("number of commands not a valid number")
+		}
+	}
+
+	// init baseVector
+	var baseVector bit.Vector
+	var vecLen uint64
+	for {
+		b, err := reader.ReadSlice('\n')
+		if err != nil && err != bufio.ErrBufferFull {
+			return fmt.Errorf("could not read bitvector: %w", err)
+		}
+
+		for _, v := range b {
+			if vecLen%bit.SubvectorBits == 0 {
+				baseVector = append(baseVector, 0)
+			}
+
+			if v == '1' {
+				baseVector.Set(vecLen)
+			}
+			vecLen++
+		}
+		if err == nil {
+			break
+		}
+	}
+
+	// Move base vector into new structure, we could also skip this step and directly read into the interleaved structure.
+	// Since this copying is not needed normally it does not contribute to the recorded runtime
+	// The precomputation of our data structure will be done later and will contribute to the runtime
+	vec := bit.NewInterleavedVectorNoPrecompute(baseVector)
+
+	// To see the implementation of the interleaved vector go to pkg/bit/interleaved_vector.go
+
+	commandFuncs := make([]CommandFunc, 0)
+
+	// scan commands
+	for {
+		line, err := reader.ReadString('\n')
+		if err == io.EOF {
+			break
+		} else if err != nil {
+			return fmt.Errorf("could not read command: %w", err)
+		}
+		line = line[:len(line)-1]
+		tokens := strings.Split(line, " ")
+
+		command := Command(tokens[0])
+		if c, ok := CommandExecutors[command]; ok {
+			args := tokens[1:]
+			commandFunc, err := c(args)
+			if err != nil {
[…]

### 2024-07-03 internal/bitvector/file_processor_test.go

@@ -1,0 +1,137 @@
+package bitvector_test
+
+import (
+	"strings"
+	"testing"
+
+	"github.com/paulheg/kit_advanced_data_structures/internal/bitvector"
+	"github.com/stretchr/testify/assert"
+)
+
+func TestFileProcessor(t *testing.T) {
+	testCases := []struct {
+		desc     string
+		input    string
+		expected string
+	}{
+		{
+			desc: "standard example",
+			input: `6
+001110110101010111111111
+access 4
+rank 0 10
+select 1 14
+rank 1 10
+select 0 3
+access 5`,
+			expected: `1
+4
+20
+6
+5
+0
+`,
+		},
+		{
+			desc: "simple test zeros",
+			input: `7
+000000000000000000000000000000000000000000000000000000000000000000000000000000000000000000
+access 0
+access 1
+access 2
+access 3
+access 4
+access 5
+access 88
+`,
+			expected: `0
+0
+0
+0
+0
+0
+0
+`,
+		},
+		{
+			desc: "simple test access",
+			input: `7
+11110000111100001111000011110000111100001111000011110000111100001111000011110000111100001111000011110000111100001111000011110000
+access 0
+access 1
+access 2
+access 3
+access 4
+access 5
+access 88
+`,
+			expected: `1
+1
+1
+1
+0
+0
+1
+`,
+		},
+		{
+			desc: "simple test access alternating",
+			input: `5
+0101010101010101010101010101010101010101010101010101010101010101010101010101010101010101010
+access 0
+access 1
+access 22
+access 23
+access 40
+`,
+			expected: `0
+1
+0
+1
+0
+`,
+		},
+		{
+			desc: "nothing working bro",
+			input: `10
[…]

### 2024-07-03 internal/bitvector/generator.go

@@ -1,0 +1,108 @@
+package bitvector
+
+import (
+	"bufio"
+	"fmt"
+	"io"
+	"math/bits"
+	"math/rand"
+	"strconv"
+	"strings"
+
+	"github.com/paulheg/kit_advanced_data_structures/pkg/bit"
+)
+
+func GenerateRandomTestCase(vectorSlices64, commands uint64, commandOut, expectedOut io.Writer) error {
+
+	commandBuffer := bufio.NewWriterSize(commandOut, 1024*1024)
+	defer commandBuffer.Flush()
+	expectedBuffer := bufio.NewWriterSize(expectedOut, 1024*1024)
+	defer expectedBuffer.Flush()
+
+	commandBuffer.Write([]byte(fmt.Sprintf("%d\n", commands)))
+
+	var ones, zeros uint64
+
+	generatorFormatString := "%0" + strconv.FormatUint(bit.SubvectorBits, 10) + "b"
+
+	var vector bit.Vector = make([]bit.Subvector, vectorSlices64)
+	for i := 0; i < int(vectorSlices64); i++ {
+		vector[i] = bit.Subvector(rand.Uint64())
+		ones += uint64(bits.OnesCount64(uint64(vector[i])))
+		binary := fmt.Sprintf(generatorFormatString, vector[i])
+
+		bBinary := []byte(binary)
+
+		for i, j := 0, len(bBinary)-1; i < j; i, j = i+1, j-1 {
+			bBinary[i], bBinary[j] = bBinary[j], bBinary[i]
+		}
+
+		n, err := commandBuffer.Write(bBinary)
+		if n != int(bit.SubvectorBits) || err != nil {
+			return fmt.Errorf("error writing, n=%d: %w", n, err)
+		}
+	}
+
+	zeros = vector.Bits() - ones
+	commandBuffer.Write([]byte{'\n'})
+
+	fullVector := bit.NewInterleavedVector(vector)
+
+	for i := 0; i < int(commands); i++ {
+		fullCommand, expectedResult, err := randomCommandAndResult(fullVector, ones, zeros)
+		if err != nil {
+			return err
+		}
+
+		commandBuffer.Write([]byte(fullCommand + "\n"))
+		expectedBuffer.Write([]byte(expectedResult + "\n"))
+	}
+
+	return nil
+}
+
+func randomCommandAndResult(vec bit.RankSelectVector, ones, zeros uint64) (fullCommand string, result string, err error) {
+
+	command, fullCommand := randomCommand(ones, zeros)
[…]

### 2024-07-03 internal/bitvector/generator_test.go

@@ -1,0 +1,18 @@
+package bitvector_test
+
+import (
+	"strings"
+	"testing"
+
+	"github.com/paulheg/kit_advanced_data_structures/internal/bitvector"
+	"github.com/stretchr/testify/assert"
+)
+
+func TestGenerator(t *testing.T) {
+
+	var commands strings.Builder
+	var expected strings.Builder
+
+	err := bitvector.GenerateRandomTestCase(10, 1000, &commands, &expected)
+	assert.NoError(t, err)
+}

### 2024-07-03 pkg/bit/bits_tables.go

@@ -1,0 +1,261 @@
+// Code generated by go run make_tables.go. DO NOT EDIT.
+
+package bit
+
+const onesLookup = "" +
+	"\x00\x00\x00\x00" +
+	"\x08\x00\x00\x00" +
+	"\x09\x00\x00\x00" +
+	"\x98\x00\x00\x00" +
+	"\x0a\x00\x00\x00" +
+	"\xa8\x00\x00\x00" +
+	"\xa9\x00\x00\x00" +
+	"\x98\x0a\x00\x00" +
+	"\x0b\x00\x00\x00" +
+	"\xb8\x00\x00\x00" +
+	"\xb9\x00\x00\x00" +
+	"\x98\x0b\x00\x00" +
+	"\xba\x00\x00\x00" +
+	"\xa8\x0b\x00\x00" +
+	"\xa9\x0b\x00\x00" +
+	"\x98\xba\x00\x00" +
+	"\x0c\x00\x00\x00" +
+	"\xc8\x00\x00\x00" +
+	"\xc9\x00\x00\x00" +
+	"\x98\x0c\x00\x00" +
+	"\xca\x00\x00\x00" +
+	"\xa8\x0c\x00\x00" +
+	"\xa9\x0c\x00\x00" +
+	"\x98\xca\x00\x00" +
+	"\xcb\x00\x00\x00" +
+	"\xb8\x0c\x00\x00" +
+	"\xb9\x0c\x00\x00" +
+	"\x98\xcb\x00\x00" +
+	"\xba\x0c\x00\x00" +
+	"\xa8\xcb\x00\x00" +
+	"\xa9\xcb\x00\x00" +
+	"\x98\xba\x0c\x00" +
+	"\x0d\x00\x00\x00" +
+	"\xd8\x00\x00\x00" +
+	"\xd9\x00\x00\x00" +
+	"\x98\x0d\x00\x00" +
+	"\xda\x00\x00\x00" +
+	"\xa8\x0d\x00\x00" +
+	"\xa9\x0d\x00\x00" +
+	"\x98\xda\x00\x00" +
+	"\xdb\x00\x00\x00" +
+	"\xb8\x0d\x00\x00" +
+	"\xb9\x0d\x00\x00" +
+	"\x98\xdb\x00\x00" +
[…]

### 2024-07-03 pkg/bit/interleaved_vector.go

@@ -1,0 +1,210 @@
+package bit
+
+import (
+	"cmp"
+	"math"
+)
+
+const InterleavedSubvectorCount uint64 = 7
+
+// Define 512 bit datastructure that will fit in one cache line
+type InterleavedVectorLine struct {
+	PreSum uint64
+	Vec    [InterleavedSubvectorCount]Subvector
+}
+
+var _ RankSelectVector = (*InterleavedVector)(nil)
+var _ Setable = (*InterleavedVector)(nil)
+
+// Create the interleaved datastructure by copying the vec Vector
+// Also create the pre sums
+func NewInterleavedVector(vec Vector) *InterleavedVector {
+	lines := uint64(math.Ceil(float64(len(vec)) / float64(InterleavedSubvectorCount)))
+
+	intlVec := &InterleavedVector{
+		vec: make([]InterleavedVectorLine, lines),
+	}
+
+	var totalSum uint64
+
+	for i := range lines {
+		startPos := i * InterleavedSubvectorCount
+		endPos := startPos + InterleavedSubvectorCount
+
+		if endPos > uint64(len(vec)) {
+			endPos = uint64(len(vec))
+		}
+
+		lineSlice := vec[startPos:endPos]
+		copy(intlVec.vec[i].Vec[:], lineSlice)
+
+		intlVec.vec[i].PreSum = totalSum
+
+		totalSum += Vector(lineSlice).Ones()
+	}
+
+	return intlVec
+}
+
+func NewInterleavedVectorNoPrecompute(vec Vector) *InterleavedVector {
+	lines := uint64(math.Ceil(float64(len(vec)) / float64(InterleavedSubvectorCount)))
+
+	intlVec := &InterleavedVector{
+		vec: make([]InterleavedVectorLine, lines),
+	}
+
+	for i := range lines {
+		startPos := i * InterleavedSubvectorCount
+		endPos := startPos + InterleavedSubvectorCount
+
+		if endPos > uint64(len(vec)) {
+			endPos = uint64(len(vec))
+		}
+
+		lineSlice := vec[startPos:endPos]
+		copy(intlVec.vec[i].Vec[:], lineSlice)
+	}
+
+	return intlVec
+}
+
+type InterleavedVector struct {
+	vec []InterleavedVectorLine
+}
+
+// Calculate the pre sums on an otherwise filled InterleavedVector
+func (i *InterleavedVector) Precompute() {
+	var sum uint64
+
+	for j := range len(i.vec) {
[…]

### 2024-07-03 pkg/bit/interleaved_vector_test.go

@@ -1,0 +1,54 @@
+package bit_test
+
+import (
+	"testing"
+
+	"github.com/paulheg/kit_advanced_data_structures/pkg/bit"
+	"github.com/stretchr/testify/assert"
+)
+
+func TestInterleavedRank(t *testing.T) {
+
+	onesVec := bit.Vector{
+		bit.Subvector(0xFFFFFFFF_FFFFFFFF), // 64
+		bit.Subvector(0xFFFFFFFF_FFFFFFFF), // 128
+		bit.Subvector(0xFFFFFFFF_FFFFFFFF), // 192
+		bit.Subvector(0xFFFFFFFF_FFFFFFFF), // 256
+		bit.Subvector(0xFFFFFFFF_FFFFFFFF), // 320
+		bit.Subvector(0xFFFFFFFF_FFFFFFFF), // 384
+		bit.Subvector(0xFFFFFFFF_FFFFFFFF), // 448
+		bit.Subvector(0xFFFFFFFF_FFFFFFFF), // 512
+		bit.Subvector(0xFFFFFFFF_FFFFFFFF), // 576
+	}
+
+	testCases := []struct {
+		desc     string
+		vec      bit.Vector
+		alpha    bool
+		position uint64
+		expected uint64
+	}{
+		{
+			desc:     "zeros in one vec - pos 266",
+			vec:      onesVec,
+			alpha:    false,
+			position: 266,
+			expected: 0,
+		},
+		{
+			desc:     "zeros in one vec - pos 522 - second line",
+			vec:      onesVec,
+			alpha:    false,
+			position: 522,
+			expected: 0,
+		},
+	}
+	for _, tC := range testCases {
+		t.Run(tC.desc, func(t *testing.T) {
+			interleaved := bit.NewInterleavedVector(tC.vec)
+			actual := interleaved.Rank(tC.alpha, tC.position)
+
+			assert.Equal(t, tC.expected, actual)
+		})
+	}
+}

### 2024-07-03 pkg/bit/make_tables.go

@@ -1,0 +1,71 @@
+//go:build ignore
+
+// This program generates bits_tables.go.
+
+package main
+
+import (
+	"bytes"
+	"fmt"
+	"go/format"
+	"io"
+	"log"
+	"os"
+)
+
+var header = []byte(`
+// Code generated by go run make_tables.go. DO NOT EDIT.
+
+package bit
+
+`)
+
+func main() {
+	buf := bytes.NewBuffer(header)
+
+	genOnes(buf)
+
+	// log.Print(buf.String())
+
+	out, err := format.Source(buf.Bytes())
+	if err != nil {
+		log.Fatal(err)
+	}
+
+	err = os.WriteFile("bits_tables.go", out, 0666)
+	if err != nil {
+		log.Fatal(err)
+	}
+}
+
+func genOnes(w io.Writer) {
+	const name = "onesLookup"
+
+	// Use a const string to allow the compiler to constant-evaluate lookups at constant index.
+	fmt.Fprintf(w, "const %s = \"\"+\n", name)
+
+	const validBit = 0b1000
+	const posMask = 0b0111
+
+	const size = 256
+
+	lookup := make([]uint32, size)
+
+	for i := 0; i < size; i++ {
+		var r uint32 = 0
+		count := 0
+		for j := 0; j < 8; j++ {
+			if i>>j&1 == 1 {
+				r |= uint32((j | validBit) << (count * 4))
+				count++
+			}
+		}
+		lookup[i] = r
+		fmt.Fprintf(w, `"`+"\\x%02x\\x%02x\\x%02x\\x%02x"+`"`, r&0xFF, (r>>8)&0xFF, (r>>16)&0xFF, (r>>24)&0xFF)
+
+		if i < size-1 {
+			w.Write([]byte("+ \n"))
+		}
+	}
+	fmt.Fprint(w, "\n\n")
+}

### 2024-07-03 pkg/bit/rankable.go

@@ -1,0 +1,11 @@
+package bit
+
+type RankableWithSize interface {
+	Rankable
+	Sizable
+}
+
+type Rankable interface {
+	// Number of alphas before position
+	Rank(alpha bool, position uint64) uint64
+}

### 2024-07-03 pkg/bit/rankable_baseline.go

@@ -1,0 +1,29 @@
+package bit
+
+var _ RankableWithSize = (*RankableBaseline)(nil)
+
+type RankableBaseline struct {
+	Vector Vector
+}
+
+// Size implements RankableWithSize.
+func (r *RankableBaseline) Size() uint64 {
+	return 0
+}
+
+// Rank implements Rankable.
+func (r *RankableBaseline) Rank(alpha bool, position uint64) uint64 {
+	var rank uint64 = 0
+
+	for i := 0; i < int(position); i++ {
+		if r.Vector.Access(uint64(i)) {
+			rank++
+		}
+	}
+
+	if alpha {
+		return rank
+	}
+
+	return position - rank
+}

### 2024-07-03 pkg/bit/rankable_test.go

@@ -1,0 +1,153 @@
+package bit_test
+
+import (
+	"fmt"
+	"math"
+	"math/rand"
+	"testing"
+
+	"github.com/paulheg/kit_advanced_data_structures/pkg/bit"
+	"github.com/stretchr/testify/assert"
+)
+
+type rankStragyBuilder func(bit.Vector) bit.Rankable
+
+var rankStrategies = map[string]rankStragyBuilder{
+	"baseline": func(v bit.Vector) bit.Rankable {
+		return &bit.RankableBaseline{v}
+	},
+	"interleaved": func(v bit.Vector) bit.Rankable {
+		return bit.NewInterleavedVector(v)
+	},
+}
+
+func convert(input []byte) bit.Vector {
+	vecLen := int(math.Ceil(float64(len(input)) / float64(bit.SubvectorBits)))
+
+	vec := make(bit.Vector, vecLen)
+
+	for i := 0; i < len(input); i++ {
+		// pos := len(input) - i - 1
+		if input[i] == 1 {
+			vec.Set(uint64(i))
+		}
+	}
+
+	return vec
+}
+
+func TestRankable(t *testing.T) {
+
+	exampleVector := convert([]byte{0, 1, 1, 0, 1, 1, 0, 1, 0, 0})
+
+	testCases := []struct {
+		desc     string
+		vec      bit.Vector
+		pos      uint64
+		alpha    bool
+		expected uint64
+	}{
+		{
+			desc:     "basic example zeros 5",
+			vec:      exampleVector,
+			pos:      5,
+			alpha:    false,
+			expected: 2,
+		},
+		{
+			desc:     "basic example zeros 3",
+			vec:      exampleVector,
+			pos:      3,
+			alpha:    false,
+			expected: 1,
+		},
+		{
+			desc:     "basic example ones",
+			vec:      exampleVector,
+			pos:      5,
+			alpha:    true,
+			expected: 3,
+		},
+		{
+			desc:     "zero position does not include current position (zeros)",
+			vec:      exampleVector,
+			pos:      0,
+			alpha:    false,
+			expected: 0,
+		},
+		{
+			desc:     "zero position does not include current position (ones)",
+			vec:      exampleVector,
+			pos:      0,
+			alpha:    true,
+			expected: 0,
+		},
+	}
+
+	for stratName, strat := range rankStrategies {
+		for _, tC := range testCases {
[…]

### 2024-07-03 pkg/bit/selectable.go

@@ -1,0 +1,11 @@
+package bit
+
+type Selectable interface {
+	// Position of the n'th alpha
+	Select(alpha bool, n uint64) uint64
+}
+
+type SelectableWithSize interface {
+	Selectable
+	Sizable
+}

### 2024-07-03 pkg/bit/selectable_baseline.go

@@ -1,0 +1,28 @@
+package bit
+
+var _ Selectable = (*SelectableBaseline)(nil)
+
+type SelectableBaseline struct {
+	Vector
+}
+
+// Select implements Selectable.
+func (s *SelectableBaseline) Select(alpha bool, n uint64) uint64 {
+	var count uint64 = 0
+
+	if n == 0 {
+		panic("n hast to be bigger than 0")
+	}
+
+	for i := uint64(0); i < s.Bits(); i++ {
+		if s.Access(i) == alpha {
+			count++
+		}
+
+		if count == n {
+			return i
+		}
+	}
+
+	panic("position not found")
+}

### 2024-07-03 pkg/bit/selectable_test.go

@@ -1,0 +1,225 @@
+package bit_test
+
+import (
+	"fmt"
+	"math/rand"
+	"testing"
+
+	"github.com/paulheg/kit_advanced_data_structures/pkg/bit"
+	"github.com/stretchr/testify/assert"
+)
+
+type selectableStrategyBulder func(vec bit.Vector) bit.Selectable
+
+var selectStrategies = map[string]selectableStrategyBulder{
+	"baseline": func(vec bit.Vector) bit.Selectable {
+		return &bit.SelectableBaseline{
+			Vector: vec,
+		}
+	},
+	"interleaved": func(vec bit.Vector) bit.Selectable {
+		return bit.NewInterleavedVector(vec)
+	},
+}
+
+func BenchmarkSelect(b *testing.B) {
+	const vecSize = 10000000
+	const operations = 100
+
+	// generate random vector
+	vector := make(bit.Vector, vecSize)
+	for i := 0; i < vecSize; i++ {
+		vector[i] = bit.Subvector(rand.Uint64())
+	}
+	ones := vector.Ones()
+
+	b.ResetTimer()
+
+	for stratName, strat := range selectStrategies {
+		b.Run(fmt.Sprintf("%s", stratName), func(b *testing.B) {
+
+			// build time
+			rv := strat(vector)
+
+			// selects
+			for i := 0; i < operations; i++ {
+				pos := uint64(rand.Int63n(int64(ones)) + 1)
+				rv.Select(true, pos)
+			}
+		})
+	}
+}
+
+func TestSelectVsNaive(t *testing.T) {
+	// generate vector
+	const size = 10_000
+	vector := make(bit.Vector, size)
+	for i := 0; i < size; i++ {
+		vector[i] = bit.Subvector(rand.Uint64())
+	}
+
+	ones := vector.Ones()
+	zeros := vector.Bits() - ones
+
+	preparedVecs := make(map[string]bit.Selectable)
+	for stratName, strat := range selectStrategies {
+		preparedVecs[stratName] = strat(vector)
+	}
+
+	naive := bit.SelectableBaseline{
+		Vector: vector,
+	}
+
+	for i := 0; i < 1000; i++ {
+		onePos := rand.Int63n(int64(ones))
+		zeroPos := rand.Int63n(int64(zeros))
+
+		expected1 := naive.Select(true, uint64(onePos))
[…]

### 2024-07-03 pkg/bit/size.go

@@ -1,0 +1,6 @@
+package bit
+
+type Sizable interface {
+	// Size in bits
+	Size() uint64
+}

### 2024-07-03 pkg/bit/vector.go

@@ -1,0 +1,178 @@
+package bit
+
+import (
+	"fmt"
+	"math"
+	"math/bits"
+)
+
+const SubvectorBits uint64 = 64
+const SubvectorMax Subvector = math.MaxUint64
+
+type Subvector uint64
+
+func (s *Subvector) Set(pos uint8) {
+	*s |= 1 << pos
+}
+
+func (s Subvector) Ones() uint8 {
+	return uint8(bits.OnesCount64(uint64(s)))
+}
+
+func (s Subvector) Access(pos uint8) bool {
+	return (s&(1<<pos))>>pos == 1
+}
+
+func (s Subvector) Rank(alpha bool, n uint8) uint8 {
+	offsetBlock := s & ^(math.MaxUint64 << n)
+	ones := uint8(bits.OnesCount64(uint64(offsetBlock)))
+	if alpha {
+		return ones
+	} else {
+		return n - ones
+	}
+}
+
+func (s Subvector) Select(alpha bool, n uint8) uint8 {
+
+	if alpha {
+		return s.OneSelect64(n)
+	} else {
+		return (^s).OneSelect64(n)
+	}
+}
+
+func (s Subvector) OneSelect64(n uint8) uint8 {
+
+	const posMask = 0b0111
+	const validMask = 0b1000
+
+	n -= 1
+
+	for i := 0; i < int(SubvectorBits); i += 8 {
+		segment := uint8(s >> i)
+		ones := uint8(bits.OnesCount8(segment))
+		if ones < n+1 {
+			n -= ones
+		} else {
+
+			// 4 bytes per number
+			var lookupBegin uint32 = uint32(segment) * 4
+			bytePos := lookupBegin + (uint32(n) >> 1)
+			valueTuple := byte(onesLookup[bytePos])
+			value := valueTuple >> ((n % 2) * 4)
+
+			isValid := value&validMask == validMask
+			if isValid {
+				// add segment offset
+				return (value & posMask) + byte(i)
+			} else {
+				panic("not found")
+			}
+		}
+	}
+
+	// should not happen
+	panic("not found")
+}
+
+var _ Accessible = (*Vector)(nil)
+var _ AccessibleWithSize = (*Vector)(nil)
+
+type Vector []Subvector
+
+// Size implements AccessibleWithSize.
+func (b *Vector) Size() uint64 {
+	return b.Bits()
+}
+
+type Accessible interface {
+	Access(position uint64) bool
+}
+
+type Setable interface {
[…]

### 2024-07-03 pkg/bit/vector_test.go

@@ -1,0 +1,300 @@
+package bit_test
+
+import (
+	"fmt"
+	"math/rand"
+	"testing"
+
+	"github.com/paulheg/kit_advanced_data_structures/pkg/bit"
+	"github.com/stretchr/testify/assert"
+)
+
+func TestBitVectorSetAndGet(t *testing.T) {
+	testCases := []struct {
+		desc     string
+		position uint64
+		size     uint64
+	}{
+		{
+			desc:     "position 0",
+			position: 0,
+			size:     1,
+		},
+		{
+			desc:     "position 1",
+			position: 1,
+			size:     1,
+		},
+		{
+			desc:     "position 63",
+			position: 63,
+			size:     1,
+		},
+		{
+			desc:     "bigger: position 64",
+			position: 64,
+			size:     2,
+		},
+	}
+	for _, tC := range testCases {
+		t.Run(tC.desc, func(t *testing.T) {
+
+			bv := make(bit.Vector, tC.size)
+
+			bv.Set(tC.position)
+			t.Logf("%b", bv)
+			assert.True(t, bv.Access(tC.position))
+
+			bv.Unset(tC.position)
+			assert.False(t, bv.Access(tC.position))
+		})
+	}
+}
+
+func TestSubvector(t *testing.T) {
+	vect := bit.Vector{0xFF00FF00FF00FF00, 0xFF00FF00FF00FF00}
+
+	testCases := []struct {
+		desc             string
+		vec              bit.Vector
+		position, length uint64
+		expected         bit.Subvector
+	}{
+		{
+			desc:     "begin",
+			position: 0,
+			length:   8,
+			expected: 0x00,
+			vec:      vect,
+		},
+		{
+			desc:     "begin with offset",
+			position: 8,
+			length:   8,
+			expected: 0xFF,
+			vec:      vect,
+		},
+		{
+			desc:     "overlap",
+			position: 60,
+			length:   8,
+			expected: 0x0F,
+			vec:      vect,
+		},
+	}
+	for _, tC := range testCases {
+		t.Run(tC.desc, func(t *testing.T) {
+
+			actual := tC.vec.Subvector(tC.position, tC.length)
+			assert.Equal(t, tC.expected, actual)
+		})
+	}
+}
+
+func TestSubvectorSelect(t *testing.T) {
+	testCases := []struct {
[…]

### 2024-07-03 article/article.tex
add link to github

@@
 \hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-% \nolinenumbers
+\nolinenumbers
 
@@
 
-\author{Paul Hegenberg}{Karlsruher Institut für Technologie, Germany \and \url{https://blog.hegenberg.dev} }{urlgh@student.kit.edu}{}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
+\author{Paul Hegenberg}{Karlsruher Institut für Technologie, Germany \and \url{https://github.com/paulheg/kit_advanced_data_structures}}{urlgh@student.kit.edu}{}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty.
 Please provide at least the name of the affiliation and the country.
@@
 \section{Algorithm}
-In the following section we will present a datastructure
+In the following section we will present a data structure
 that implements \code{rank}, \code{select} and \code{access} operations.
@@
 
+\section{Code}
+The code for this implementation can be found on Github in the following repository: 
+\url{https://github.com/paulheg/kit_advanced_data_structures}.
+
 \end{document}
